---
title: "HNB DigiMedia Data Exploration"
subtitle: "Gathering information for map design"
author: "HNB DigiMedia Project"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    theme: cosmo
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(data.table)

hnb <- readRDS("C:/Users/lsikic/Dropbox/hnb.rds")
setDT(hnb)
```

# 1. Basic Data Structure

## 1.1 Dimensions and Variables

```{r dimensions}
cat("Total articles:", nrow(hnb), "\n")
cat("Variables:", ncol(hnb), "\n")
cat("\nVariable names:\n")
names(hnb)
```

## 1.2 Date Range and Temporal Coverage

```{r date_range}
hnb[, date := as.Date(DATE)]

cat("Date range:", as.character(min(hnb$date, na.rm = TRUE)), 
    "to", as.character(max(hnb$date, na.rm = TRUE)), "\n")

hnb[, .(n_articles = .N), by = year(date)][order(year)] |> print()
```

```{r monthly_coverage}
monthly <- hnb[, .(n = .N), by = .(year = year(date), month = month(date))]
monthly[, plot_date := as.Date(paste(year, month, "01", sep = "-"))]

ggplot(monthly, aes(x = plot_date, y = n)) +

geom_line(linewidth = 0.8, color = "#2c3e50") +
geom_vline(xintercept = as.Date("2020-07-10"), linetype = "dashed", 
           color = "#e74c3c", alpha = 0.7) +
geom_vline(xintercept = as.Date("2023-01-01"), linetype = "dashed", 
           color = "#27ae60", alpha = 0.7) +
annotate("text", x = as.Date("2020-07-10"), y = max(monthly$n) * 0.95, 
         label = "ERM II", hjust = -0.1, color = "#e74c3c", size = 3) +
annotate("text", x = as.Date("2023-01-01"), y = max(monthly$n) * 0.95, 
         label = "Euro", hjust = -0.1, color = "#27ae60", size = 3) +
labs(title = "Monthly Article Volume",
     x = NULL, y = "Number of Articles") +
theme_minimal()
```

# 2. Source Analysis

## 2.1 Source Distribution

```{r source_count}
n_sources <- uniqueN(hnb$FROM)
cat("Total unique sources:", n_sources, "\n")
```

```{r top_sources}
source_counts <- hnb[, .(
n_articles = .N,
pct = .N / nrow(hnb) * 100
), by = FROM][order(-n_articles)]

cat("\nTop 30 sources:\n")
source_counts[1:30] |> print()
```

```{r source_concentration}
source_counts[, cumulative_pct := cumsum(pct)]

cat("\nConcentration analysis:\n")
cat("Top 5 sources cover:", round(source_counts[5, cumulative_pct], 1), "% of articles\n")
cat("Top 10 sources cover:", round(source_counts[10, cumulative_pct], 1), "% of articles\n")
cat("Top 20 sources cover:", round(source_counts[20, cumulative_pct], 1), "% of articles\n")

sources_for_80 <- source_counts[cumulative_pct <= 80, .N] + 1
cat("Sources needed for 80% coverage:", sources_for_80, "\n")
```

```{r source_distribution_plot}
ggplot(source_counts[1:30], aes(x = reorder(FROM, n_articles), y = n_articles)) +
geom_col(fill = "#3498db", alpha = 0.8) +
coord_flip() +
labs(title = "Top 30 Sources by Article Count",
     x = NULL, y = "Number of Articles") +
theme_minimal() +
scale_y_continuous(labels = comma)
```

## 2.2 Full Source List for Classification

```{r all_sources}
cat("All", nrow(source_counts), "sources with article counts:\n\n")
print(source_counts[, .(FROM, n_articles, pct = round(pct, 2))], nrows = Inf)
```

## 2.3 Domain Patterns from URL

```{r url_patterns}
hnb[, domain := str_extract(URL, "(?<=://)[^/]+")]
hnb[, domain := str_remove(domain, "^www\\.")]

domain_counts <- hnb[, .(n = .N), by = domain][order(-n)]

cat("Top 30 domains:\n")
domain_counts[1:30] |> print()
```

# 3. REACH Analysis

## 3.1 REACH Distribution

```{r reach_overview}
cat("REACH variable summary:\n")
summary(hnb$REACH)

cat("\nNon-zero REACH:", sum(hnb$REACH > 0, na.rm = TRUE), 
    "(", round(mean(hnb$REACH > 0, na.rm = TRUE) * 100, 1), "%)\n")
cat("Missing REACH:", sum(is.na(hnb$REACH)), "\n")
```

```{r reach_distribution}
hnb_reach <- hnb[REACH > 0]

ggplot(hnb_reach, aes(x = REACH)) +
geom_histogram(bins = 50, fill = "#2ecc71", alpha = 0.7) +
scale_x_log10(labels = comma) +
labs(title = "Distribution of REACH (log scale)",
     subtitle = paste0("N = ", nrow(hnb_reach), " articles with REACH > 0"),
     x = "REACH", y = "Count") +
theme_minimal()
```

## 3.2 REACH by Source

```{r reach_by_source}
source_reach <- hnb[, .(
n_articles = .N,
total_reach = sum(REACH, na.rm = TRUE),
mean_reach = mean(REACH, na.rm = TRUE),
median_reach = median(REACH, na.rm = TRUE)
), by = FROM][order(-total_reach)]

cat("Top 30 sources by total REACH:\n")
source_reach[1:30] |> print()
```

```{r reach_concentration}
source_reach[, reach_pct := total_reach / sum(total_reach) * 100]
source_reach[, reach_cumulative := cumsum(reach_pct)]

cat("\nREACH concentration:\n")
cat("Top 5 sources capture:", round(source_reach[5, reach_cumulative], 1), "% of total REACH\n")
cat("Top 10 sources capture:", round(source_reach[10, reach_cumulative], 1), "% of total REACH\n")
cat("Top 20 sources capture:", round(source_reach[20, reach_cumulative], 1), "% of total REACH\n")
```

## 3.3 Article Count vs REACH Comparison

```{r count_vs_reach}
comparison <- merge(
source_counts[, .(FROM, rank_articles = 1:.N)],
source_reach[, .(FROM, rank_reach = 1:.N)],
by = "FROM"
)

comparison[, rank_diff := rank_articles - rank_reach]

cat("Sources with biggest rank differences (article count vs REACH):\n")
cat("\nHigher REACH rank than article rank (punching above weight):\n")
comparison[order(-rank_diff)][1:10] |> print()

cat("\nLower REACH rank than article rank (punching below weight):\n")
comparison[order(rank_diff)][1:10] |> print()
```

# 4. Author Analysis

```{r author_overview}
cat("Articles with author:", sum(!is.na(hnb$AUTHOR) & hnb$AUTHOR != ""), "\n")
cat("Unique authors:", uniqueN(hnb$AUTHOR[!is.na(hnb$AUTHOR) & hnb$AUTHOR != ""]), "\n")

author_counts <- hnb[!is.na(AUTHOR) & AUTHOR != "", .(n = .N), by = AUTHOR][order(-n)]

cat("\nTop 30 authors:\n")
author_counts[1:30] |> print()
```

# 5. Content Analysis Preparation

## 5.1 Title and Text Availability

```{r text_availability}
cat("TITLE availability:", sum(!is.na(hnb$TITLE) & hnb$TITLE != ""), "/", nrow(hnb), "\n")
cat("FULL_TEXT availability:", sum(!is.na(hnb$FULL_TEXT) & hnb$FULL_TEXT != ""), "/", nrow(hnb), "\n")

hnb[, text_length := nchar(FULL_TEXT)]
cat("\nText length (characters):\n")
summary(hnb$text_length)
```

## 5.2 Sample Titles

```{r sample_titles}
cat("Random sample of 30 titles:\n\n")
set.seed(42)
sample(hnb$TITLE[!is.na(hnb$TITLE)], 30) |> cat(sep = "\n\n")
```

## 5.3 Keyword Frequency in Titles

```{r keyword_check}
keywords <- c("HNB", "Hrvatska narodna banka", "guverner", "Vujčić", 
            "inflacija", "euro", "kuna", "tečaj", "kamat", 
            "ECB", "monetarn", "deviz", "banka")

keyword_freq <- sapply(keywords, function(kw) {
sum(str_detect(tolower(hnb$TITLE), tolower(kw)), na.rm = TRUE)
})

data.frame(
keyword = keywords,
n_articles = keyword_freq,
pct = round(keyword_freq / nrow(hnb) * 100, 1)
) |> arrange(desc(n_articles)) |> print()
```

## 5.4 Pronadjena_recenica Field

This appears to be the sentence that matched your search criteria.

```{r matched_sentence}
cat("Sample of matched sentences:\n\n")
set.seed(123)
sample(hnb$Pronadjena_recenica[!is.na(hnb$Pronadjena_recenica)], 10) |> 
cat(sep = "\n\n---\n\n")
```

# 6. Temporal Patterns

## 6.1 Day of Week Distribution

```{r dow_distribution}
hnb[, dow := lubridate::wday(date, label = TRUE)]

dow_counts <- hnb[, .(n = .N), by = dow]

ggplot(dow_counts, aes(x = dow, y = n)) +
geom_col(fill = "#9b59b6", alpha = 0.8) +
labs(title = "Articles by Day of Week",
     x = NULL, y = "Count") +
theme_minimal()
```

## 6.2 Articles by Policy Period

```{r period_analysis}
hnb[, period := fcase(
date < as.Date("2020-07-10"), "Pre-ERM II",
date < as.Date("2023-01-01"), "ERM II",
default = "Euro Era"
)]

hnb[, period := factor(period, levels = c("Pre-ERM II", "ERM II", "Euro Era"))]

period_summary <- hnb[, .(
n_articles = .N,
n_sources = uniqueN(FROM),
total_reach = sum(REACH, na.rm = TRUE),
mean_reach = mean(REACH, na.rm = TRUE),
days = as.numeric(max(date) - min(date))
), by = period]

period_summary[, articles_per_day := round(n_articles / days, 1)]

print(period_summary)
```

## 6.3 Top Sources by Period

```{r sources_by_period}
for (p in c("Pre-ERM II", "ERM II", "Euro Era")) {
cat("\n===", p, "===\n")
hnb[period == p, .N, by = FROM][order(-N)][1:10] |> print()
}
```

## 6.4 REACH Distribution by Period

```{r reach_by_period}
ggplot(hnb[REACH > 0], aes(x = period, y = REACH)) +
geom_boxplot(fill = "#3498db", alpha = 0.6) +
scale_y_log10(labels = comma) +
labs(title = "REACH Distribution by Period",
     x = NULL, y = "REACH (log scale)") +
theme_minimal()
```

# 7. Classification Signals

## 7.1 Financial Media Indicators

```{r classification_signals}
financial_terms <- c("poslovni", "lider", "finance", "ekonom", "business", 
                   "novac", "kapital", "banka", "seebiz", "money", "tportal")

source_counts[, has_financial_term := str_detect(tolower(FROM), 
                                                  paste(financial_terms, collapse = "|"))]

cat("Sources with financial terms in name:\n")
source_counts[has_financial_term == TRUE, .(FROM, n_articles)] |> print()

cat("\nTheir combined share:", 
  round(sum(source_counts[has_financial_term == TRUE, n_articles]) / nrow(hnb) * 100, 1), 
  "% of articles\n")
```

## 7.2 General News Portal Patterns

```{r news_portal_patterns}
news_terms <- c("index", "jutarnji", "večernji", "24sata", "dnevnik", 
              "net.hr", "telegram", "nacional", "slobodna", "novi list")

source_counts[, has_news_term := str_detect(tolower(FROM), 
                                             paste(news_terms, collapse = "|"))]

cat("Sources matching general news patterns:\n")
source_counts[has_news_term == TRUE, .(FROM, n_articles)] |> print()
```

## 7.3 Potential HNB Official Sources

```{r hnb_sources}
hnb_terms <- c("hnb", "narodna banka")

source_counts[, is_hnb := str_detect(tolower(FROM), paste(hnb_terms, collapse = "|"))]

cat("Potential HNB official sources:\n")
source_counts[is_hnb == TRUE, .(FROM, n_articles)] |> print()
```

# 8. KEYWORD_NAME Analysis

```{r keyword_name}
cat("Distribution of KEYWORD_NAME:\n")
hnb[, .N, by = KEYWORD_NAME][order(-N)] |> print()
```

# 9. Data Quality Summary

```{r quality_summary}
key_vars <- c("DATE", "FROM", "URL", "TITLE", "FULL_TEXT", "REACH", "AUTHOR")

quality_check <- data.frame(
Variable = key_vars,
Non_Missing = sapply(key_vars, function(v) {
  sum(!is.na(hnb[[v]]) & hnb[[v]] != "")
}),
Pct_Complete = sapply(key_vars, function(v) {
  round(mean(!is.na(hnb[[v]]) & hnb[[v]] != "") * 100, 1)
})
)

print(quality_check)
```

# 10. Export for Classification

```{r export_sources}
source_export <- merge(source_counts, source_reach[, .(FROM, total_reach, mean_reach)], 
                      by = "FROM", all.x = TRUE)

source_export <- source_export[order(-n_articles), .(
FROM, 
n_articles, 
pct_articles = round(pct, 2),
total_reach,
mean_reach = round(mean_reach, 0),
actor_category = ""
)]

fwrite(source_export, "hnb_source_classification.csv")
cat("Exported hnb_source_classification.csv with", nrow(source_export), "sources\n")
```

---

# Summary Statistics for Map Design

Copy these key numbers after running:

```{r final_summary}
cat("="
, rep("=", 50), "\n", sep = "")
cat("HNB DIGIMEDIA DATA SUMMARY\n")
cat(rep("=", 51), "\n\n", sep = "")

cat("Total articles:", nrow(hnb), "\n")
cat("Date range:", as.character(min(hnb$date)), "to", as.character(max(hnb$date)), "\n")
cat("Unique sources:", uniqueN(hnb$FROM), "\n")
cat("Unique authors:", uniqueN(hnb$AUTHOR[!is.na(hnb$AUTHOR) & hnb$AUTHOR != ""]), "\n\n")

cat("REACH coverage:", round(mean(hnb$REACH > 0, na.rm = TRUE) * 100, 1), "% non-zero\n")
cat("Total REACH:", format(sum(hnb$REACH, na.rm = TRUE), big.mark = ","), "\n\n")

cat("Articles by period:\n")
print(hnb[, .N, by = period])

cat("\nTop 5 sources:\n")
print(source_counts[1:5, .(FROM, n_articles, pct = round(pct, 1))])

cat("\nTop 5 sources by REACH:\n")
print(source_reach[1:5, .(FROM, total_reach, pct = round(reach_pct, 1))])
```
