---
title: "HNB DigiMedia Map 6: Source Concentration and Market Structure"
subtitle: "Analyzing Attention Market Concentration and Gatekeeper Vulnerability"
author: "HNB DigiMedia Project"
date: today
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    fig-width: 10
    fig-height: 6
execute:
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| include: false

library(tidyverse)
library(data.table)
library(scales)
library(ggplot2)
library(knitr)
library(kableExtra)
library(viridis)
library(lubridate)
library(patchwork)
library(ggrepel)
library(ineq)  # For Gini coefficient

theme_set(theme_minimal(base_size = 12) +
          theme(
            plot.title = element_text(face = "bold", size = 14),
            plot.subtitle = element_text(color = "gray40"),
            legend.position = "bottom"
          ))

# Color palette for actor types
actor_colors <- c(
  "HNB Official" = "#1a3c5a",
  "Financial Regulators" = "#2c5f7c",
  "Commercial Banks" = "#4a9c6d",
  "Financial Media" = "#e07b39",
  "General Media" = "#c44536",
  "International Sources" = "#6b4c9a",
  "Academic/Expert" = "#64748b",
  "Political Actors" = "#eab308",
  "Financial Influencers" = "#ec4899",
  "Government" = "#0ea5e9",
  "Other" = "#94a3b8"
)

# Period colors
period_colors <- c(
  "ERM II" = "#3498db",
  "Euro Era" = "#27ae60"
)

# Concentration thresholds (DOJ/FTC guidelines adapted)
hhi_thresholds <- c(
  "Unconcentrated" = 1500,
  "Moderately Concentrated" = 2500,
  "Highly Concentrated" = 10000
)
```

# Introduction

This document presents **Map 6: Source Concentration and Market Structure Analysis** of the HNB DigiMedia project. The core question driving this analysis is:

> **How concentrated is the attention market for HNB coverage, and does this concentration create systemic vulnerabilities for monetary policy communication?**

## Theoretical Motivation

The literature establishes that monetary policy works primarily through managing expectations (Woodford 2001). If a handful of media sources dominate public attention on central bank matters, then:

1. **Gatekeeper dependency:** HNB communication effectiveness depends critically on how these gatekeepers frame content
2. **Fragility risk:** Changes in gatekeeper behavior (editorial shifts, ownership changes, platform decline) could disrupt communication channels
3. **Message distortion:** Concentrated intermediation increases opportunities for framing divergence from official messaging

Morris and Shin (2002) showed that public information can crowd out private information when it receives "too much weight." In a concentrated attention market, dominant sources become de facto public signals that may override diverse interpretations.

## Market Structure Framework

We apply industrial organization concepts to the attention market:

| Metric | Definition | Interpretation |
|--------|------------|----------------|
| Gini Coefficient | Inequality measure (0-1) | Higher = more concentrated |
| HHI | Sum of squared market shares | >2500 = highly concentrated |
| CR4 / CR10 | Top 4/10 concentration ratio | Share of top sources |
| Power Law α | Zipf distribution exponent | Steeper = more inequality |

```{r load-data}
#| label: load-data

hnb <- readRDS("C:/Users/lsikic/Dropbox/hnb.rds")
setDT(hnb)

hnb[, date := as.Date(DATE)]
hnb[, year := year(date)]
hnb[, month := month(date)]

hnb[, period := fcase(
  date < as.Date("2023-01-01"), "ERM II",
  default = "Euro Era"
)]
hnb[, period := factor(period, levels = c("ERM II", "Euro Era"))]

n_articles <- nrow(hnb)
n_sources <- uniqueN(hnb$FROM)
total_reach <- sum(hnb$REACH, na.rm = TRUE)

cat("Data loaded:", n_articles, "articles from", n_sources, "sources\n")
```

```{r actor-classification}
#| label: actor-classification

# =============================================================================
# ACTOR CLASSIFICATION - SYNCHRONIZED WITH MAPS 1-5
# =============================================================================

manual_overrides <- list(
  "HNB Official" = c(
    "hnb.hr", "hrvatska narodna banka",
    "hrvatska narodna banka / croatian national bank"
  ),
  "Financial Regulators" = c(
    "hanfa.hr", "hanfa", "hrvatska agencija za nadzor financijskih usluga",
    "fina.hr", "fina", "financijska agencija"
  ),
  "Government" = c(
    "gov.hr", "vlada", "vlada republike hrvatske",
    "ministarstvo financija", "sabor.hr"
  ),
  "Financial Media" = c(
    "poslovni.hr", "poslovni dnevnik", "seebiz.eu", "seebiz",
    "lider.media", "lidermedia.hr", "lider", "bloombergadria.com",
    "bloomberg adria", "hrportfolio.hr", "privredni.hr"
  ),
  "Commercial Banks" = c(
    "zaba.hr", "zagrebačka banka", "pbz.hr", "privredna banka zagreb",
    "erste.hr", "erste banka", "hpb.hr", "hrvatska poštanska banka",
    "otp banka", "raiffeisen", "addiko"
  ),
  "Academic/Expert" = c(
    "eizg.hr", "ekonomski institut zagreb", "ijf.hr",
    "institut za javne financije", "ekonomski fakultet"
  )
)

general_media_sources <- c(
  "index.hr", "index", "jutarnji.hr", "jutarnji", "jutarnji list",
  "vecernji.hr", "vecernji", "večernji list",
  "24sata.hr", "24sata", "slobodnadalmacija.hr", "slobodna dalmacija",
  "novilist.hr", "novi list", "dnevnik.hr", "dnevnik",
  "tportal.hr", "tportal", "telegram.hr", "telegram",
  "net.hr", "rtl.hr", "rtl", "hrt.hr", "hrt",
  "n1info.hr", "n1info.com", "n1", "nacional.hr", "nacional",
  "direktno.hr", "direktno", "dnevno.hr", "dnevno"
)

classify_actor_hnb <- function(from_val, url_val = NA) {
  from_lower <- tolower(trimws(as.character(from_val)))
  url_lower <- tolower(ifelse(is.na(url_val), "", as.character(url_val)))
  combined <- paste(from_lower, url_lower)
  
  match_any <- function(patterns, text, fixed = TRUE) {
    any(sapply(patterns, function(p) grepl(p, text, fixed = fixed)))
  }
  
  for (actor_type in names(manual_overrides)) {
    if (match_any(manual_overrides[[actor_type]], from_lower)) {
      return(actor_type)
    }
  }
  
  if (grepl("hnb\\.hr", url_lower) || grepl("hnb\\.hr", from_lower)) {
    return("HNB Official")
  }
  
  if (grepl("hanfa|fina\\.hr", combined)) {
    return("Financial Regulators")
  }
  
  if (grepl("gov\\.hr|vlada|ministarstvo", combined)) {
    return("Government")
  }
  
  bank_patterns <- c("banka", "bank", "zaba", "pbz", "erste", "hpb",
                     "otp", "raiffeisen", "addiko", "sberbank")
  if (match_any(bank_patterns, from_lower)) {
    if (!grepl("narodna|hnb|central", from_lower)) {
      return("Commercial Banks")
    }
  }
  
  financial_media_patterns <- c(
    "poslovni", "seebiz", "lider", "bloomberg", "portfolio",
    "privredni", "finance", "ekonom", "business", "capital"
  )
  if (match_any(financial_media_patterns, from_lower)) {
    return("Financial Media")
  }
  
  if (match_any(general_media_sources, from_lower)) {
    return("General Media")
  }
  
  news_patterns <- c("portal", "vijesti", "news", "info", "dnevni",
                     "radio", "tv", ".hr")
  if (match_any(news_patterns, from_lower)) {
    return("General Media")
  }
  
  return("Other")
}

hnb[, ACTOR_TYPE := mapply(classify_actor_hnb, FROM, URL)]
hnb[, INSTITUTIONAL := ACTOR_TYPE %in% c("HNB Official", "Financial Regulators", "Government")]
```

# Analysis 6.1: Overall Market Concentration

## Source-Level Market Shares

```{r source-shares}
#| label: source-shares

source_stats <- hnb[, .(
  Articles = .N,
  Total_REACH = sum(REACH, na.rm = TRUE),
  Mean_REACH = mean(REACH, na.rm = TRUE),
  ACTOR_TYPE = first(ACTOR_TYPE),
  Institutional = first(INSTITUTIONAL)
), by = FROM][order(-Total_REACH)]

source_stats[, `:=`(
  Article_Share = Articles / sum(Articles) * 100,
  REACH_Share = Total_REACH / sum(Total_REACH) * 100,
  Rank = .I
)]

source_stats[, Cumulative_REACH_Share := cumsum(REACH_Share)]
source_stats[, Cumulative_Article_Share := cumsum(Article_Share)]
```

## Concentration Metrics Summary

```{r concentration-metrics}
#| label: concentration-metrics

# Gini coefficients
gini_reach <- ineq(source_stats$Total_REACH, type = "Gini")
gini_articles <- ineq(source_stats$Articles, type = "Gini")

# HHI (Herfindahl-Hirschman Index) - using percentage shares squared, summed
hhi_reach <- sum((source_stats$REACH_Share)^2)
hhi_articles <- sum((source_stats$Article_Share)^2)

# Concentration ratios
cr1_reach <- source_stats[1, REACH_Share]
cr4_reach <- sum(source_stats[1:min(4, nrow(source_stats)), REACH_Share])
cr10_reach <- sum(source_stats[1:min(10, nrow(source_stats)), REACH_Share])
cr20_reach <- sum(source_stats[1:min(20, nrow(source_stats)), REACH_Share])

cr1_articles <- source_stats[1, Article_Share]
cr4_articles <- sum(source_stats[1:min(4, nrow(source_stats)), Article_Share])
cr10_articles <- sum(source_stats[1:min(10, nrow(source_stats)), Article_Share])
cr20_articles <- sum(source_stats[1:min(20, nrow(source_stats)), Article_Share])

# HHI classification
hhi_class <- case_when(
  hhi_reach < 1500 ~ "Unconcentrated",
  hhi_reach < 2500 ~ "Moderately Concentrated",
  TRUE ~ "Highly Concentrated"
)

tibble(
  Metric = c(
    "Number of Sources",
    "Gini Coefficient (REACH)",
    "Gini Coefficient (Articles)",
    "HHI (REACH)",
    "HHI (Articles)",
    "HHI Classification",
    "CR1 (Top Source REACH Share)",
    "CR4 (Top 4 REACH Share)",
    "CR10 (Top 10 REACH Share)",
    "CR20 (Top 20 REACH Share)"
  ),
  Value = c(
    format(n_sources, big.mark = ","),
    sprintf("%.3f", gini_reach),
    sprintf("%.3f", gini_articles),
    sprintf("%.0f", hhi_reach),
    sprintf("%.0f", hhi_articles),
    hhi_class,
    sprintf("%.1f%%", cr1_reach),
    sprintf("%.1f%%", cr4_reach),
    sprintf("%.1f%%", cr10_reach),
    sprintf("%.1f%%", cr20_reach)
  ),
  Interpretation = c(
    "Total unique publishers in corpus",
    "0 = perfect equality, 1 = perfect inequality",
    "0 = perfect equality, 1 = perfect inequality",
    "<1500 unconcentrated, >2500 highly concentrated",
    "<1500 unconcentrated, >2500 highly concentrated",
    "Based on DOJ/FTC market concentration guidelines",
    "Single source dominance",
    "Oligopoly threshold (>60% = tight oligopoly)",
    "Extended market concentration",
    "Broad market concentration"
  )
) %>%
  kable(col.names = c("Metric", "Value", "Interpretation")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Lorenz Curve: REACH Inequality

```{r lorenz-curve}
#| label: lorenz-curve
#| fig-height: 7

# Create Lorenz curve data
lorenz_data <- source_stats[order(Total_REACH), .(
  FROM, Total_REACH, Articles
)]
lorenz_data[, `:=`(
  cum_sources = .I / .N * 100,
  cum_reach = cumsum(Total_REACH) / sum(Total_REACH) * 100,
  cum_articles = cumsum(Articles) / sum(Articles) * 100
)]

# Add origin point
lorenz_plot <- rbind(
  data.table(cum_sources = 0, cum_reach = 0, cum_articles = 0),
  lorenz_data[, .(cum_sources, cum_reach, cum_articles)]
)

ggplot(lorenz_plot) +
  geom_line(aes(x = cum_sources, y = cum_sources), 
            linetype = "dashed", color = "gray50") +
  geom_line(aes(x = cum_sources, y = cum_reach), 
            color = "#e07b39", linewidth = 1.2) +
  geom_line(aes(x = cum_sources, y = cum_articles), 
            color = "#2c5f7c", linewidth = 1.2) +
  geom_ribbon(aes(x = cum_sources, ymin = cum_reach, ymax = cum_sources),
              fill = "#e07b39", alpha = 0.2) +
  annotate("text", x = 70, y = 30, 
           label = paste0("Gini (REACH) = ", sprintf("%.3f", gini_reach)),
           color = "#e07b39", fontface = "bold", size = 4) +
  annotate("text", x = 70, y = 45, 
           label = paste0("Gini (Articles) = ", sprintf("%.3f", gini_articles)),
           color = "#2c5f7c", fontface = "bold", size = 4) +
  labs(
    title = "Lorenz Curve: Attention Inequality in HNB Coverage",
    subtitle = "Orange = REACH distribution, Blue = Article distribution, Dashed = Perfect equality",
    x = "Cumulative % of Sources (ranked low to high)",
    y = "Cumulative % of Total",
    caption = "The further the curve from the diagonal, the greater the concentration"
  ) +
  coord_equal() +
  scale_x_continuous(limits = c(0, 100)) +
  scale_y_continuous(limits = c(0, 100))
```

## Cumulative Distribution

```{r cumulative-distribution}
#| label: cumulative-distribution
#| fig-height: 6

# How many sources needed for X% of REACH
reach_50 <- source_stats[Cumulative_REACH_Share >= 50, .N][1]
reach_80 <- source_stats[Cumulative_REACH_Share >= 80, .N][1]
reach_90 <- source_stats[Cumulative_REACH_Share >= 90, .N][1]

ggplot(source_stats[1:min(100, nrow(source_stats))], 
       aes(x = Rank, y = Cumulative_REACH_Share)) +
  geom_line(color = "#e07b39", linewidth = 1.2) +
  geom_hline(yintercept = c(50, 80, 90), linetype = "dashed", color = "gray50", alpha = 0.7) +
  geom_vline(xintercept = c(reach_50, reach_80, reach_90), 
             linetype = "dotted", color = "#2c5f7c", alpha = 0.7) +
  annotate("text", x = reach_50 + 2, y = 52, 
           label = paste0(reach_50, " sources = 50%"), hjust = 0, size = 3) +
  annotate("text", x = reach_80 + 2, y = 82, 
           label = paste0(reach_80, " sources = 80%"), hjust = 0, size = 3) +
  annotate("text", x = reach_90 + 2, y = 92, 
           label = paste0(reach_90, " sources = 90%"), hjust = 0, size = 3) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    title = "Cumulative REACH Distribution",
    subtitle = "How many sources account for X% of total attention?",
    x = "Number of Sources (ranked by REACH)",
    y = "Cumulative REACH Share (%)"
  )
```

# Analysis 6.2: Attention Gatekeepers

## Top 20 Sources by REACH

```{r top-sources}
#| label: top-sources
#| fig-height: 9

top20 <- source_stats[1:20]

ggplot(top20, aes(x = reorder(FROM, Total_REACH), y = Total_REACH, fill = ACTOR_TYPE)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = sprintf("%.1f%%", REACH_Share)), hjust = -0.1, size = 3) +
  coord_flip() +
  scale_fill_manual(values = actor_colors) +
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = "Top 20 Attention Gatekeepers",
    subtitle = "Sources with highest total REACH in HNB coverage",
    x = NULL,
    y = "Total REACH",
    fill = "Actor Type"
  )
```

## Gatekeeper Profile Table

```{r gatekeeper-table}
#| label: gatekeeper-table

top20 %>%
  mutate(
    Articles = format(Articles, big.mark = ","),
    Total_REACH = format(Total_REACH, big.mark = ","),
    Mean_REACH = format(round(Mean_REACH, 0), big.mark = ","),
    Article_Share = sprintf("%.2f%%", Article_Share),
    REACH_Share = sprintf("%.2f%%", REACH_Share),
    Cumulative_REACH_Share = sprintf("%.1f%%", Cumulative_REACH_Share),
    Institutional = ifelse(Institutional, "Yes", "No")
  ) %>%
  select(Rank, FROM, ACTOR_TYPE, Institutional, Articles, Article_Share, 
         REACH_Share, Cumulative_REACH_Share, Mean_REACH) %>%
  kable(col.names = c("Rank", "Source", "Actor Type", "Inst.", "Articles", 
                      "Article %", "REACH %", "Cumul. %", "Mean REACH")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  scroll_box(height = "500px")
```

## Gatekeeper Concentration by Actor Type

```{r gatekeeper-actor}
#| label: gatekeeper-actor
#| fig-height: 6

# What actor types dominate the top 20?
top20_actor <- top20[, .(
  Sources = .N,
  Total_REACH = sum(Total_REACH),
  REACH_Share = sum(REACH_Share)
), by = ACTOR_TYPE][order(-REACH_Share)]

ggplot(top20_actor, aes(x = reorder(ACTOR_TYPE, REACH_Share), y = REACH_Share, fill = ACTOR_TYPE)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = sprintf("%.1f%% (%d)", REACH_Share, Sources)), hjust = -0.1, size = 3.5) +
  coord_flip() +
  scale_fill_manual(values = actor_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  labs(
    title = "Actor Type Composition of Top 20 Gatekeepers",
    subtitle = "Which actor types control the attention gates?",
    x = NULL,
    y = "Combined REACH Share (%)",
    caption = "Numbers show REACH share and count of sources in top 20"
  ) +
  theme(legend.position = "none")
```

# Analysis 6.3: Power Law Analysis

Media attention markets often follow power law distributions where a few sources dominate while most have minimal reach. We test whether HNB coverage follows this pattern.

## Rank-Size Distribution (Zipf Plot)

```{r zipf-plot}
#| label: zipf-plot
#| fig-height: 7

# Log-log plot for power law detection
source_stats[, log_rank := log10(Rank)]
source_stats[, log_reach := log10(Total_REACH + 1)]

# Fit linear model on log-log scale
zipf_model <- lm(log_reach ~ log_rank, data = source_stats[Total_REACH > 0])
zipf_alpha <- -coef(zipf_model)[2]  # Zipf exponent (negative of slope)
zipf_r2 <- summary(zipf_model)$r.squared

ggplot(source_stats[Total_REACH > 0], aes(x = log_rank, y = log_reach)) +
  geom_point(alpha = 0.5, color = "#2c5f7c") +
  geom_smooth(method = "lm", se = TRUE, color = "#e07b39", fill = "#e07b39", alpha = 0.2) +
  annotate("text", x = max(source_stats$log_rank) * 0.7, y = max(source_stats$log_reach) * 0.9,
           label = paste0("Zipf α = ", sprintf("%.2f", zipf_alpha), "\nR² = ", sprintf("%.3f", zipf_r2)),
           hjust = 0, size = 4, fontface = "bold") +
  labs(
    title = "Power Law Test: Rank-Size Distribution",
    subtitle = "Log-log plot tests whether REACH follows Zipf's law",
    x = "Log10(Rank)",
    y = "Log10(REACH)",
    caption = "Linear relationship on log-log scale indicates power law; α > 1 indicates high concentration"
  )
```

## Power Law Interpretation

```{r power-law-interpretation}
#| label: power-law-interpretation

cat("Power Law Analysis Results\n")
cat("==========================\n")
cat("Zipf exponent (α):", sprintf("%.2f", zipf_alpha), "\n")
cat("R-squared:", sprintf("%.3f", zipf_r2), "\n")
cat("\nInterpretation:\n")
if (zipf_r2 > 0.8) {
  cat("- Strong power law fit (R² > 0.8)\n")
} else if (zipf_r2 > 0.6) {
  cat("- Moderate power law fit (R² 0.6-0.8)\n")
} else {
  cat("- Weak power law fit (R² < 0.6)\n")
}

if (zipf_alpha > 1.5) {
  cat("- Very high concentration (α > 1.5): Winner-take-most dynamics\n")
} else if (zipf_alpha > 1) {
  cat("- High concentration (α > 1): Top sources dominate significantly\n")
} else {
  cat("- Moderate concentration (α ≤ 1): More distributed attention\n")
}

cat("\nImplication for HNB communication:\n")
if (zipf_alpha > 1 && zipf_r2 > 0.7) {
  cat("The attention market follows predictable power law dynamics.\n")
  cat("A small number of gatekeepers will consistently dominate reach.\n")
  cat("Strategic relationships with top sources are critical.\n")
}
```

# Analysis 6.4: Period Comparison

## Concentration Changes: ERM II vs Euro Era

```{r period-concentration}
#| label: period-concentration

# Calculate concentration metrics by period
period_concentration <- hnb[, {
  source_period <- .SD[, .(Total_REACH = sum(REACH, na.rm = TRUE)), by = FROM][order(-Total_REACH)]
  source_period[, REACH_Share := Total_REACH / sum(Total_REACH) * 100]
  
  list(
    Sources = uniqueN(FROM),
    Gini = ineq(source_period$Total_REACH, type = "Gini"),
    HHI = sum(source_period$REACH_Share^2),
    CR4 = sum(source_period$REACH_Share[1:min(4, .N)]),
    CR10 = sum(source_period$REACH_Share[1:min(10, .N)]),
    Top_Source = source_period$FROM[1],
    Top_Source_Share = source_period$REACH_Share[1]
  )
}, by = period]

period_concentration %>%
  mutate(
    Sources = format(Sources, big.mark = ","),
    Gini = sprintf("%.3f", Gini),
    HHI = sprintf("%.0f", HHI),
    CR4 = sprintf("%.1f%%", CR4),
    CR10 = sprintf("%.1f%%", CR10),
    Top_Source_Share = sprintf("%.1f%%", Top_Source_Share)
  ) %>%
  kable(col.names = c("Period", "Sources", "Gini", "HHI", "CR4", "CR10", 
                      "Top Source", "Top Source %")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Concentration Trend Visualization

```{r concentration-trend}
#| label: concentration-trend
#| fig-height: 6

period_conc_long <- period_concentration %>%
  select(period, Gini, CR4, CR10) %>%
  mutate(CR4 = CR4 / 100, CR10 = CR10 / 100) %>%  # Normalize to 0-1 scale
  pivot_longer(cols = -period, names_to = "Metric", values_to = "Value")

ggplot(period_conc_long, aes(x = period, y = Value, fill = Metric)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.2f", Value)), 
            position = position_dodge(width = 0.6), vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = c("Gini" = "#1a3c5a", "CR4" = "#e07b39", "CR10" = "#4a9c6d")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)), limits = c(0, 1)) +
  labs(
    title = "Concentration Metrics by Period",
    subtitle = "Did euro adoption change market concentration?",
    x = "Period",
    y = "Concentration Index (0-1 scale)",
    fill = "Metric"
  ) +
  theme(legend.position = "top")
```

# Analysis 6.5: Gatekeeper Stability

## Do the Same Sources Dominate Across Periods?

```{r gatekeeper-stability}
#| label: gatekeeper-stability

# Get top 20 sources in each period
top_erm2 <- hnb[period == "ERM II", .(Total_REACH = sum(REACH, na.rm = TRUE)), by = FROM][order(-Total_REACH)][1:20]
top_euro <- hnb[period == "Euro Era", .(Total_REACH = sum(REACH, na.rm = TRUE)), by = FROM][order(-Total_REACH)][1:20]

# Calculate overlap
overlap_sources <- intersect(top_erm2$FROM, top_euro$FROM)
overlap_pct <- length(overlap_sources) / 20 * 100

# Rank correlation
both_periods <- merge(
  top_erm2[, .(FROM, REACH_ERM2 = Total_REACH, Rank_ERM2 = .I)],
  top_euro[, .(FROM, REACH_Euro = Total_REACH, Rank_Euro = .I)],
  by = "FROM", all = TRUE
)

# Spearman correlation for overlapping sources
overlap_data <- both_periods[!is.na(Rank_ERM2) & !is.na(Rank_Euro)]
if (nrow(overlap_data) >= 3) {
  rank_cor <- cor(overlap_data$Rank_ERM2, overlap_data$Rank_Euro, method = "spearman")
} else {
  rank_cor <- NA
}

cat("Gatekeeper Stability Analysis\n")
cat("=============================\n")
cat("Top 20 overlap between periods:", length(overlap_sources), "sources\n")
cat("Overlap percentage:", sprintf("%.0f%%", overlap_pct), "\n")
cat("Rank correlation (Spearman):", ifelse(is.na(rank_cor), "N/A", sprintf("%.2f", rank_cor)), "\n")
cat("\nOverlapping gatekeepers:\n")
for (s in overlap_sources) {
  cat("  -", s, "\n")
}
```

## Gatekeeper Movement Visualization

```{r gatekeeper-movement}
#| label: gatekeeper-movement
#| fig-height: 8

# Prepare data for bump chart style visualization
movement_data <- merge(
  top_erm2[, .(FROM, Rank_ERM2 = .I)],
  top_euro[, .(FROM, Rank_Euro = .I)],
  by = "FROM", all = TRUE
)
movement_data[is.na(Rank_ERM2), Rank_ERM2 := 25]  # New entrants
movement_data[is.na(Rank_Euro), Rank_Euro := 25]   # Dropped out

movement_data[, Movement := Rank_ERM2 - Rank_Euro]  # Positive = improved
movement_data[, Status := fcase(
  Rank_ERM2 == 25, "New in Euro Era",
  Rank_Euro == 25, "Dropped from Top 20",
  abs(Movement) <= 3, "Stable",
  Movement > 0, "Improved",
  Movement < 0, "Declined"
)]

# Show sources with biggest movements
movement_data[order(-abs(Movement))][1:15] %>%
  mutate(
    Rank_ERM2 = ifelse(Rank_ERM2 == 25, "Not in Top 20", as.character(Rank_ERM2)),
    Rank_Euro = ifelse(Rank_Euro == 25, "Not in Top 20", as.character(Rank_Euro)),
    Movement = sprintf("%+d", Movement)
  ) %>%
  kable(col.names = c("Source", "Rank (ERM II)", "Rank (Euro Era)", "Movement", "Status"),
        caption = "Gatekeeper Position Changes") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Analysis 6.6: Fragility Assessment

## What If Top Gatekeepers Shift Coverage?

This analysis quantifies the fragility of the attention ecosystem. If top sources changed their HNB coverage behavior, how much attention would be affected?

```{r fragility-analysis}
#| label: fragility-analysis
#| fig-height: 7

# Calculate concentration at different thresholds
fragility_data <- data.table(
  Top_N = c(1, 3, 5, 10, 20, 50),
  REACH_Share = sapply(c(1, 3, 5, 10, 20, 50), function(n) {
    sum(source_stats$REACH_Share[1:min(n, nrow(source_stats))])
  }),
  Article_Share = sapply(c(1, 3, 5, 10, 20, 50), function(n) {
    sum(source_stats$Article_Share[1:min(n, nrow(source_stats))])
  })
)

fragility_data[, Sources_Pct := Top_N / n_sources * 100]

fragility_long <- fragility_data %>%
  pivot_longer(cols = c(REACH_Share, Article_Share), 
               names_to = "Metric", values_to = "Share") %>%
  mutate(Metric = ifelse(Metric == "REACH_Share", "REACH", "Articles"))

ggplot(fragility_long, aes(x = factor(Top_N), y = Share, fill = Metric)) +
  geom_col(position = "dodge", width = 0.7) +
  geom_text(aes(label = sprintf("%.0f%%", Share)), 
            position = position_dodge(width = 0.7), vjust = -0.3, size = 3) +
  scale_fill_manual(values = c("REACH" = "#e07b39", "Articles" = "#2c5f7c")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)), limits = c(0, 100)) +
  labs(
    title = "Fragility Analysis: Concentration at Different Thresholds",
    subtitle = "What share do the top N sources control?",
    x = "Number of Top Sources",
    y = "Share of Total (%)",
    fill = "Metric"
  ) +
  theme(legend.position = "top")
```

## Fragility Summary Table

```{r fragility-table}
#| label: fragility-table

fragility_data %>%
  mutate(
    `% of All Sources` = sprintf("%.1f%%", Sources_Pct),
    `REACH Share` = sprintf("%.1f%%", REACH_Share),
    `Article Share` = sprintf("%.1f%%", Article_Share),
    `Concentration Ratio` = sprintf("%.1fx", REACH_Share / Article_Share)
  ) %>%
  select(Top_N, `% of All Sources`, `Article Share`, `REACH Share`, `Concentration Ratio`) %>%
  kable(col.names = c("Top N Sources", "% of All Sources", "Article Share", 
                      "REACH Share", "REACH/Article Ratio"),
        caption = "How concentrated is attention at different thresholds?") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## HNB Position in Market Structure

```{r hnb-position}
#| label: hnb-position

# Where does HNB Official rank?
hnb_rank <- source_stats[ACTOR_TYPE == "HNB Official", Rank][1]
hnb_reach_share <- source_stats[ACTOR_TYPE == "HNB Official", REACH_Share][1]
hnb_article_share <- source_stats[ACTOR_TYPE == "HNB Official", Article_Share][1]

# How many non-institutional sources rank above HNB?
sources_above_hnb <- source_stats[Rank < hnb_rank]
media_above_hnb <- sources_above_hnb[ACTOR_TYPE %in% c("Financial Media", "General Media")]

cat("HNB Official Position in Market Structure\n")
cat("==========================================\n")
cat("HNB Official rank:", hnb_rank, "of", n_sources, "sources\n")
cat("HNB REACH share:", sprintf("%.2f%%", hnb_reach_share), "\n")
cat("HNB Article share:", sprintf("%.2f%%", hnb_article_share), "\n")
cat("\nSources ranking above HNB Official:", nrow(sources_above_hnb), "\n")
cat("  - Media sources above HNB:", nrow(media_above_hnb), "\n")
cat("\nThese", nrow(sources_above_hnb), "sources capture",
    sprintf("%.1f%%", sum(sources_above_hnb$REACH_Share)), "of total REACH\n")
```

# Summary and Key Findings

## Market Structure Summary

```{r final-summary}
#| label: final-summary

tibble(
  Finding = c(
    "Total sources in ecosystem",
    "Gini coefficient (REACH)",
    "HHI classification",
    "Top source REACH share",
    "Top 4 sources REACH share",
    "Top 10 sources REACH share",
    "Sources needed for 50% REACH",
    "Sources needed for 80% REACH",
    "Power law exponent (α)",
    "Gatekeeper stability (overlap %)",
    "HNB Official rank",
    "HNB Official REACH share"
  ),
  Value = c(
    format(n_sources, big.mark = ","),
    sprintf("%.3f", gini_reach),
    hhi_class,
    sprintf("%.1f%%", cr1_reach),
    sprintf("%.1f%%", cr4_reach),
    sprintf("%.1f%%", cr10_reach),
    as.character(reach_50),
    as.character(reach_80),
    sprintf("%.2f", zipf_alpha),
    sprintf("%.0f%%", overlap_pct),
    as.character(hnb_rank),
    sprintf("%.2f%%", hnb_reach_share)
  )
) %>%
  kable(col.names = c("Finding", "Value")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Implications for HNB Communication Strategy

Based on the concentration analysis:

**1. Gatekeeper Dependency**

The `r sprintf("%.1f%%", cr4_reach)` CR4 ratio indicates that `r ifelse(cr4_reach > 60, "a tight oligopoly", ifelse(cr4_reach > 40, "a loose oligopoly", "a competitive market"))` controls HNB coverage attention. This means:

- Strategic relationships with top `r min(4, reach_50)` sources are critical
- Editorial decisions at these outlets disproportionately affect HNB message reach
- Diversification of media relationships reduces single-point-of-failure risk

**2. Fragility Risk**

With `r reach_50` sources controlling 50% of attention:

- The attention market is `r ifelse(reach_50 < 5, "highly fragile", ifelse(reach_50 < 10, "moderately fragile", "relatively resilient"))`
- Ownership changes, editorial shifts, or platform decline at top sources could significantly disrupt communication channels
- Building redundancy through multiple high-reach relationships is advisable

**3. Power Law Dynamics**

The Zipf exponent of `r sprintf("%.2f", zipf_alpha)` indicates:

- `r ifelse(zipf_alpha > 1, "Winner-take-most dynamics where top sources will consistently dominate", "More distributed attention patterns")`
- Marginal sources have minimal impact regardless of content quality
- Resources should focus on high-reach channels rather than broad distribution

**4. Structural Position**

HNB Official ranks #`r hnb_rank` with `r sprintf("%.2f%%", hnb_reach_share)` REACH share, meaning:

- `r nrow(sources_above_hnb)` sources capture more attention than official HNB communication
- Direct communication competes with `r sprintf("%.0f", nrow(media_above_hnb))` media sources for visibility
- The attention gap (Map 5) is structurally embedded in market concentration

---

# Appendix: Methodological Notes

## Concentration Metrics

**Gini Coefficient:** Measures inequality on a 0-1 scale. For n sources with shares $x_i$:
$$G = \frac{\sum_{i=1}^{n}\sum_{j=1}^{n}|x_i - x_j|}{2n\sum_{i=1}^{n}x_i}$$

**HHI (Herfindahl-Hirschman Index):** Sum of squared percentage market shares:
$$HHI = \sum_{i=1}^{n}s_i^2$$

Where $s_i$ is the percentage market share of source $i$. Range: 0 to 10,000.

**Concentration Ratios (CR):** Sum of market shares for top N sources.

**Zipf Exponent:** From regression of log(REACH) on log(Rank): $\log(REACH) = \alpha - \beta \log(Rank)$

## DOJ/FTC Guidelines (adapted)

- HHI < 1,500: Unconcentrated market
- HHI 1,500-2,500: Moderately concentrated
- HHI > 2,500: Highly concentrated

```{r session-info}
#| label: session-info

sessionInfo()
```
